---
title: "STA437 Final Project"
date: "2025-03-16"
output: html_document
---

# {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


cdsdata <- get(load("~/Desktop/Fourth Year/STA437/CDS_data.RData"))
head(cdsdata)

```


## 1. Introduction {.tabset}

### Motivation:
Why is this dataset interesting for you? What makes it suitable for multivariate analysis


### Research Question: 
What are you trying to uncover in the data? If you follow a guiding question, state it here. Guiding Questions for Analysis

Risk Patterns & Market Conditions
How do CDS spreads behave in calm vs. crisis periods (e.g., during COVID-19)?
Are some industries or companies consistently seen as high-risk?
Multivariate Relationships & Dependencies
How do CDS spreads for different maturities (short-term vs. long-term) move together for the same company?
Can we use Principal Component Analysis (PCA) or Canonical Correlation Analysis (CCA) to find key risk factors driving CDS spreads?
Clustering & Risk Groups
Can we identify clusters of companies that have similar risk trends? (e.g., using hierarchical clustering or k-means)
Do companies within the same industry tend to have similar CDS spread behavior?
Advanced Dependency Analysis
How do time-varying dependencies between companies evolve? (e.g., using copulas to measure joint risk behavior)
Can we model the spillover effect—how risk changes in one company affect others?




### Brief Summary of Methods: 
Mention which multivariate methods you expect to use and why.
Background (Optional): If applicable, briefly mention any relevant literature, prior knowledge, or similar analyses.

## 2. Data and Preprocessing {.tabset}

### Description of the Dataset: 
Provide basic summary statistics and visualizations of the dataset.

```{r}

```


### Include summary tables and histograms/boxplots where relevant.

### Handling of Missing Data: 
Describe any missing values and the steps taken to address them.

### Transformations:
Mention if any transformations (e.g., normalization, log-scaling) were applied.

## 3. Methodology {.tabset}

Describe the multivariate methods used and justify their relevance to your research question.

## 4. Results {.tabset}

### Data Summaries

Present key findings from visualizations and descriptive statistics.

### Interpretation of Findings.

### Conclusion
How do the results answer your research question? Were the results expected or surprising?

