---
title: "STA437 Final Project"
date: "2025-03-16"
output: html_document
---

# {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


cdsdata <- get(load("~/Desktop/Fourth Year/STA437/CDS_data.RData"))
head(cdsdata)

```


mridata <- get(load("~/Desktop/Fourth Year/STA437/ABIDE_YALE.RData"))
head(mri)



## 1. Introduction {.tabset}

### Motivation:
Why is this dataset interesting for you? What makes it suitable for multivariate analysis


### Research Question: 
What are you trying to uncover in the data? If you follow a guiding question, state it here. Guiding Questions for Analysis

Risk Patterns & Market Conditions
How do CDS spreads behave in calm vs. crisis periods (e.g., during COVID-19)?
Are some industries or companies consistently seen as high-risk?
Multivariate Relationships & Dependencies
How do CDS spreads for different maturities (short-term vs. long-term) move together for the same company?
Can we use Principal Component Analysis (PCA) or Canonical Correlation Analysis (CCA) to find key risk factors driving CDS spreads?
Clustering & Risk Groups
Can we identify clusters of companies that have similar risk trends? (e.g., using hierarchical clustering or k-means)
Do companies within the same industry tend to have similar CDS spread behavior?
Advanced Dependency Analysis
How do time-varying dependencies between companies evolve? (e.g., using copulas to measure joint risk behavior)
Can we model the spillover effect—how risk changes in one company affect others?




### Brief Summary of Methods: 
Mention which multivariate methods you expect to use and why.
Background (Optional): If applicable, briefly mention any relevant literature, prior knowledge, or similar analyses.

## 2. Data and Preprocessing {.tabset}

### Description of the Dataset: 
Provide basic summary statistics and visualizations of the dataset.

```{r}
# Number of unique companies
num_companies <- length(unique(cdsdata$Company))
cat("Number of unique companies:", num_companies, "\n")

# Summary statistics for PX1 through PX10
px_columns <- paste0("PX", 1:10)
summary_stats <- summary(cdsdata[, px_columns])
print(summary_stats)

# Standard deviation for PX1 through PX10
sd_values <- sapply(cdsdata[, px_columns], sd, na.rm = TRUE)
sd_values

# Histograms for PX1 through PX10
library(ggplot2)
library(tidyr)

cdsdata_long <- pivot_longer(cdsdata, cols = all_of(px_columns), names_to = "PX", values_to = "Value")

ggplot(cdsdata_long, aes(x = log(Value))) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  facet_wrap(~PX, scales = "free") +
  theme_minimal() +
  labs(title = "Histograms for PX1 to PX10")

```

50 companies with greatest variance in PX1

```{r}
# Number of unique companies
num_companies <- length(unique(cdsdata$Company))
cat("Number of unique companies:", num_companies, "\n")

# Summary statistics for PX1 through PX10
px_columns <- paste0("PX", 1:10)
summary_stats <- summary(cdsdata[, px_columns])
print(summary_stats)

# Standard deviation for PX1 through PX10
sd_values <- sapply(cdsdata[, px_columns], sd, na.rm = TRUE)
sd_values

# Calculate variance of PX1 for each company
company_variance <- tapply(cdsdata$PX1, cdsdata$Company, sd, na.rm = TRUE)

# Get the top 50 companies with the greatest variance in PX1
top_50_companies <- names(sort(company_variance, decreasing = TRUE))[1:10]

# Filter the original dataset to include only these 50 companies
filtered_data <- cdsdata[cdsdata$Company %in% top_50_companies, ]

# Histograms for PX1 through PX10 for the top 50 companies
library(ggplot2)
library(tidyr)

# Reshape the filtered data for plotting
filtered_data_long <- pivot_longer(filtered_data, cols = all_of(px_columns), names_to = "PX", values_to = "Value")

# Plot histograms
ggplot(filtered_data_long, aes(x = log(Value))) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  facet_wrap(~PX, scales = "free") +
  theme_minimal() +
  labs(title = "Histograms for PX1 to PX10 (Top 50 Companies with Greatest Variance in PX1)")

```


### Include summary tables and histograms/boxplots where relevant.

### Handling of Missing Data: 
Describe any missing values and the steps taken to address them.

### Transformations:
Mention if any transformations (e.g., normalization, log-scaling) were applied.

## 3. Methodology {.tabset}

Describe the multivariate methods used and justify their relevance to your research question.

```{r}
# Standardizing the CDS data for clustering
px_data <- filtered_data[, paste0("PX", 1:10)]
px_scaled <- scale(px_data)

# Calculating the distance matrix (Euclidean distance)
dist_matrix <- dist(px_scaled, method = "euclidean")

# Performing hierarchical clustering
hclust_result <- hclust(dist_matrix, method = "ward.D2")

# Plotting the dendrogram
plot(hclust_result, main = "Dendrogram of Companies", xlab = "Company", ylab = "Distance")

```

```{r}
# Standardizing the CDS data for clustering
px_data <- filtered_data[, paste0("PX", 1:10)]
px_scaled <- scale(px_data)

# Finding the optimal number of clusters using the elbow method
set.seed(123)  # For reproducibility
wss <- sapply(1:10, function(k) {
  kmeans(px_scaled, centers = k, nstart = 25)$tot.withinss
})

# Plotting the elbow method
library(ggplot2)
elbow_data <- data.frame(k = 1:10, WSS = wss)
ggplot(elbow_data, aes(x = k, y = WSS)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Elbow Method for Optimal Number of Clusters", x = "Number of Clusters", y = "Within-Cluster Sum of Squares")

# Applying K-means clustering with the chosen number of clusters (e.g., k = 3)
set.seed(123)
kmeans_result <- kmeans(px_scaled, centers = 3, nstart = 25)

# Adding cluster labels to the dataset
filtered_data$Cluster <- kmeans_result$cluster

# Visualizing the clusters (e.g., using the first two principal components)
library(ggplot2)
pca_result <- prcomp(px_scaled)
pca_data <- data.frame(PC1 = pca_result$x[, 1], PC2 = pca_result$x[, 2], Cluster = factor(filtered_data$Cluster))

ggplot(pca_data, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point() +
  theme_minimal() +
  labs(title = "K-means Clustering of Companies", x = "Principal Component 1", y = "Principal Component 2")

```


## 4. Results {.tabset}

### Data Summaries

Present key findings from visualizations and descriptive statistics.

### Interpretation of Findings.

### Conclusion
How do the results answer your research question? Were the results expected or surprising?

