---
title: "STA437 Final Project"
date: "2025-03-26"
output: html_document
---

# {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(lme4)        # For mixed-effects models
library(ggplot2)     # For visualization
library(reshape2)    # For matrix reshaping
library(igraph)      # For network analysis
library(corrplot)    # For correlation visualization

load("~/Desktop/Fourth Year/STA437/ABIDE_YALE.RData")


dem_mri <- YALE_demo_var


```

```{r}
# Compute summary by DX_GROUP
demographic_table <- dem_mri %>%
  group_by(DX_GROUP) %>%
  summarise(
    N = n(),
    `Mean Age (SD)` = sprintf("%.2f (%.2f)", mean(AGE_AT_SCAN, na.rm = TRUE), sd(AGE_AT_SCAN, na.rm = TRUE)),
    `Male (%)` = sprintf("%d (%.1f%%)", sum(SEX == "1"), 100 * mean(SEX == "1")),
    `Female (%)` = sprintf("%d (%.1f%%)", sum(SEX == "2"), 100 * mean(SEX == "2"))
  ) %>%
  rename(`Diagnosis Group` = DX_GROUP)

# Compute overall summary
overall_summary <- dem_mri %>%
  summarise(
    `Diagnosis Group` = "Overall",
    N = n(),
    `Mean Age (SD)` = sprintf("%.2f (%.2f)", mean(AGE_AT_SCAN, na.rm = TRUE), sd(AGE_AT_SCAN, na.rm = TRUE)),
    `Male (%)` = sprintf("%d (%.1f%%)", sum(SEX == "1"), 100 * mean(SEX == "1")),
    `Female (%)` = sprintf("%d (%.1f%%)", sum(SEX == "2"), 100 * mean(SEX == "2"))
  )

# Combine overall row with the grouped summary
final_table <- bind_rows(demographic_table, overall_summary)

# Print table using kable
kable(final_table, caption = "Demographic Characteristics", align = "c") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```


```{r}
subject_ids <- names(YALE_fmri)

head(YALE_fmri)
subject_ids
```

```{r}
# Compute FC for each subject
compute_fc <- function(time_series) {
  cor(time_series)  # Pearson correlation matrix (110 x 110)
}

# Apply function to all subjects
fc_matrices <- lapply(YALE_fmri, compute_fc)

# Get subject IDs (assuming order matches demographics)
subject_ids <- names(YALE_fmri)

# Convert FC matrices to long format
fc_list <- lapply(1:length(fc_matrices), function(i) {
  mat <- fc_matrices[[i]]
  df <- as.data.frame(as.table(mat))  # Convert matrix to long format
  df$SubjectID <- subject_ids[i]      # Add subject ID
  return(df)
})

# Combine all subjects into a single dataframe
fc_data <- do.call(rbind, fc_list)
colnames(fc_data) <- c("Region1", "Region2", "FC", "SubjectID")

# Merge with demographics
fc_data <- merge(fc_data, dem_mri, by = "SubjectID")

```

```{r}
# Get subject IDs
subject_ids <- seq_along(fc_matrices)  # If subject names aren't available, use indices

# Convert FC matrices to long format and add SubjectID
fc_list <- lapply(1:length(fc_matrices), function(i) {
  mat <- fc_matrices[[i]]
  df <- as.data.frame(as.table(mat))  # Convert matrix to long format
  df$SubjectID <- subject_ids[i]      # Add subject ID column
  return(df)
})

# Combine all subjects into a single dataframe
fc_data <- do.call(rbind, fc_list)

# Rename columns correctly
colnames(fc_data) <- c("Region1", "Region2", "FC", "SubjectID")

# Check output
head(fc_data)

```

```{r}
dem_mri$SubjectID <- 1:nrow(dem_mri)

fc_data <- merge(fc_data, dem_mri, by = "SubjectID")

library(lme4)
lmm <- lmer(FC ~ DX_GROUP + AGE_AT_SCAN + SEX + (1 | SubjectID), data = fc_data)
summary(lmm)

```



```{r}
library(corrplot)

# Compute group-average FC matrices
avg_fc_autism <- Reduce("+", fc_matrices[dem_mri$DX_GROUP == 1]) / sum(dem_mri$DX_GROUP == 1)
avg_fc_control <- Reduce("+", fc_matrices[dem_mri$DX_GROUP == 2]) / sum(dem_mri$DX_GROUP == 2)

par(mfrow = c(1, 2))
corrplot(avg_fc_autism, method = "color", title = "Autism Group", mar = c(0,0,1,0))
corrplot(avg_fc_control, method = "color", title = "Control Group", mar = c(0,0,1,0))

```



## 1. Introduction {.tabset}

### Motivation:
Why is this dataset interesting for you? What makes it suitable for multivariate analysis


### Research Question: 
What are you trying to uncover in the data? If you follow a guiding question, state it here. Guiding Questions for Analysis


Understanding Brain Connectivity: 
How can we detect patterns of brain connectivity from these fMRI recordings? 
Can we identify distinct functional networks in the brain using this data? 

Comparing Groups (Autism vs. Control): 
Do brain connectivity patterns differ between individuals with Autism and the Control group? 
Are there specific brain regions that show different activity between groups? 
Exploring Demographic Factors: Does age influence brain connectivity patterns? 
Are there any differences in connectivity between males and females? 

Finding Unique Relationships: 
Are there connectivity patterns that are specific to individuals with Autism but not present in the Control group? 
Can we use this data to predict diagnosis based on brain activity alone?



### Brief Summary of Methods: 
Mention which multivariate methods you expect to use and why.
Background (Optional): If applicable, briefly mention any relevant literature, prior knowledge, or similar analyses.

## 2. Data and Preprocessing {.tabset}

### Description of the Dataset: 
Provide basic summary statistics and visualizations of the dataset.


### Include summary tables and histograms/boxplots where relevant.

### Handling of Missing Data: 
Describe any missing values and the steps taken to address them.

### Transformations:
Mention if any transformations (e.g., normalization, log-scaling) were applied.

## 3. Methodology {.tabset}

Describe the multivariate methods used and justify their relevance to your research question.


## 4. Results {.tabset}

### Data Summaries

Present key findings from visualizations and descriptive statistics.

### Interpretation of Findings.

### Conclusion
How do the results answer your research question? Were the results expected or surprising?

